---
title: "R Objects, Workflow and Functions"
format: html
editor: visual
---

# Vectors

Create a vector

```{r}
set.seed(12345)
my_runif <- runif(30)
print(my_runif)
is.vector(my_runif)
```

Subset that thing

```{r}
my_runif[1:10]
my_runif[c(1,3,14:15)]
```

Sort this thing

```{r}
sort(my_runif)
```

Character vector

```{r}
char_vec <- c("23a","1b","sdc","abd")
sort(char_vec)
```
# Data frames

```{r}
data(trees)
trees
str(trees)
```
Subset a column

```{r}
trees$Girth
```
Get attributes 

```{r}
attributes(trees)
names(trees)
colnames(trees)[2:3]
```
# Lists

investiagitng our data frame

```{r}
is.list(trees)
is.data.frame(trees)
```

```{r}
trees[1:2]
trees[[1]]
```
look at the ln model

```{r}
fit <- lm(Volume ~ Height | Girth, data=trees)
```

Look at the structure using restric function()

```{r}
str(fit, max.level = 1)

```

Some helper functions

```{r}
fit$coefficients
coef(fit)

fit$residuals
residuals(fit)

#rank(fit)
```

# if/then/else

Fizz buzz challenge

-take a number
-if it is divisable by 3 return fizz
if its divisable by 5 teturn buzz
-if it is divisable by 15 return fizz buzz


```{r}
number <- 2

if((number %% 15) == 0) {
  print("fizz buzz")
} else if((number %% 5) == 0){
  print("buzz")
} else if((number %% 3) == 0){
  print("fizz")
} else {
  print("error")
}

```

